---
title: "Alone Analysis"
author: "Rahim Jutha"
date: "2023-02-01"
output: html_document
---

# Introduction

https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-24/readme.md

# Load Packages

```{r}
library(tidyverse) # Data manipulation
library(showtext) # Adding fonts
library(alone) # Alone TV Show
library(skimr) # Provides skim function
library(ggbump) # Proivdes ggbump function to make bump plots
```

# Download the Data
```{r}
episodes <- alone::episodes
loadouts <- alone::loadouts
seasons <- alone::seasons
contestants <- alone::survivalists
```

# Quick EDA

There are 4 different datasets related to the Alone Series.
- Episode Information
- Season Information
- Survivalist (Contestants) Data
- Loadout Data

```{r}
episodes %>%
  skim()
# air data information
# title
# ratings and viewership information

# Ratings for each season 
# some missing information for ratings
```

```{r}
seasons %>%
  skim()
# 9 Seasons in total
# Each season is filmed in a different location
# Date info only for the last few observations
# contains lat and long

# could maybe do a map with with the location, seasons, and dates?
# could impute the missing dates
```

```{r}
survivalists %>%
  skim()
```

```{r}
loadouts %>%
  skim()

# item vs item detailed
# most picked item each season

# or proprtiono of top 10 picked each season
```

# Visualization

https://github.com/KittJonathan/tidytuesday/blob/master/R/2023_04_alone.R

https://twitter.com/KittJonathan/status/1618288297360621568/photo/1

Going to use this visualization as a guide.

Since there are 4 different datasets lets make a graph for each and combine the 4

# Preprocessing

```{r}
# Since there are multiple seasons filmed in a location we combine the like information for those seasons
map_labels <- seasons %>%
  arrange(location, season)  %>%
  mutate(site = paste(location, country, sep = " - ")) %>%
  select(site, lat, lon, season) %>%
  group_by(site, lat, lon) %>%
  summarise(seasons = str_c(season, collapse = ",")) %>%
  rowid_to_column() # To make calling the seasons easier in ggplot
  
map_labels
```

```{r}
ratings <- episodes %>%
  group_by(season) %>%
  summarise(
    min_rating = min(imdb_rating, na.rm = TRUE),
    avg_rating = mean(imdb_rating, na.rm = TRUE),
    max_rating = max(imdb_rating, na.rm = TRUE)
  ) %>%
  mutate(avg_rating = round(avg_rating, digits = 1),
         y.pos = 9:1) # for the plot so we can have S1 at the top
```

```{r}
segment_vals <- survivalists %>% 
  filter(result == 1) %>%
  select(season, days_lasted) %>%
  distinct(season, .keep_all = TRUE) %>%
  pull(days_lasted) %>%
  lapply(function(days_lasted) seq(0, days_lasted, 1)) %>%
  unlist() %>%
  tibble(day = .) %>%
  mutate(y.pos = rep(9:1, survivalists %>% filter(result == 1) %>% distinct(season, .keep_all = TRUE) %>% pull(days_lasted) + 1))

all_days <- survivalists %>%
  select(days_lasted, season) %>%
  mutate(y.pos = case_when(
    season == 1 ~ 9,
    season == 2 ~ 8,
    season == 3 ~ 7,
    season == 4 ~ 6,
    season == 5 ~ 5,
    season == 6 ~ 4,
    season == 7 ~ 3,
    season == 8 ~ 2,
    season == 9 ~ 1,
  ))
```

```{r}
items <- 
  loadouts %>%
  select(season, item) %>%
  group_by(season, item) %>%
  mutate(count = n()) %>%
  arrange(season, -count, item) %>%
  distinct() %>%
  group_by(season) %>%
  mutate(order = season) %>%
  group_by(order) %>%
  mutate(rank = rank(-count, ties.method = "first")) %>%
  ungroup() %>%
  group_by(item) %>%
  mutate(any_top_5 = any(rank <= 5)) %>% 
  ungroup() %>%
  mutate(rank = if_else(rank > 5, 6L, rank)) %>%
  filter(any_top_5 == TRUE) %>%
  group_by(item) %>%
  mutate(first_top5 = min(order[rank <= 5]),
         last_top5 = max(order[rank <= 5]),
         d_first_top5 = if_else(order == first_top5, 1, 0)) %>%
  filter(!is.na(first_top5),
         order >= first_top5,
         order <= last_top5) %>%
  ungroup() %>%
  arrange(item, order) %>% 
  group_by(item) %>% 
  mutate(lag_zero = if_else(lag(rank) %in% c(6, NA) & rank <= 5, 1, 0, 0)) %>% 
  ungroup() %>% 
  mutate(group = cumsum(lag_zero)) %>% 
  select(item, season, order, rank, first_top5, last_top5, d_first_top5, group) %>%
  mutate(item = case_when(
    item == "Bow and arrows" ~ "Bow & arrows",
    TRUE ~ item
  ))


```

# Import Fonts

```{r}
font_add_google(name = "Roboto Condensed", family = "Roboto Condensed")
font_add_google(name = "Bebas Neue", family = "Bebas Neue")
showtext_auto()
```

```{r}

```
# Seasons World Map Plot

https://eriqande.github.io/rep-res-web/lectures/making-maps-with-R.html

```{r}
# function for adding labels and their boxes
# xmin and xmax give the x dimensions of the box,
# ymax and ymin give the y dimensions of the box,
# id tells us which label and coordinates to use,
# xend and yend give the coordinates for the line,
# text coordinates are calculated using the midpoints of box dimensions
label_box <- function(xmin, xmax, ymin, ymax, id, xend, yend){
  list(
    # top rectangle
    geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymax - 7, ymax = ymax), fill = "#3E4C59", col = "#3E4C59", linewidth = 0.25),
    # bottom rectangle
    geom_rect(aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymin + 7), fill = "#F0F0F2", col = "#3E4C59", linewidth = 0.25),
    # top text
    geom_text(data = filter(map_labels, rowid == id), aes(x = mean(c(xmin,xmax)), y = mean(c(ymax, ymax - 7)), label = site), hjust = 0.5, vjust = 0.5, col = "#F0F0F2", family = "Roboto Condensed", size = 8),
    # bottom text
    geom_text(data = filter(map_labels, rowid == id), aes(x = mean(c(xmax, xmin)), y = mean(c(ymin, ymin + 7)), label = paste0("Season ", seasons)), hjust = 0.5, vjust = 0.5, col = "#3E4C59", family = "Roboto Condensed", size = 8),
    # line 
    geom_segment(data = filter(map_labels, rowid == id), aes(x = lon, xend = xend, y = lat, yend = yend), color = "#3E4C59", linewidth = 0.25)
  )
}

# get world map and exclude Antarctica
world <- map_data("world") %>%
  filter(region != "Antarctica")

m <- ggplot() + 
  # World Map
  geom_polygon(data = world, aes(x = long, y = lat, group = group), fill = "#C5D3D9", colour = "#F0F0F2", lwd = .3) +
  coord_fixed(ratio = 1.3) +
  geom_point(data = map_labels, aes(x = lon, y = lat), col = "#3E4C59") +
  # Theme
  theme_void() +
  theme(panel.background = element_rect(fill = "#F0F0F2", color = NA)) +
  label_box(xmin = -120, xmax = -40, ymin = 34, ymax = 48, id = 1, xend = -120, yend = 48) +
  label_box(xmin = -220, xmax = -120, ymin = 71, ymax = 85, id = 2, xend = -120, yend = 71) +
  label_box(xmin = -60, xmax = 20, ymin = 65, ymax = 79, id = 3, xend = -60, yend = 65) +
  label_box(xmin = -175, xmax = -95, ymin = -37, ymax = -23, id = 4, xend = -95, yend = -37) +
  label_box(xmin = -200, xmax = -130, ymin = 26, ymax = 40, id = 5, xend = -130, yend = 40) +
  label_box(xmin = 110, xmax = 210, ymin = 28, ymax = 42, id = 6, xend = 110, yend = 42) 
```

Episode Ratings

```{r}
add_point_text <- function(column, point_fill, text_col){
  list(
      geom_point(data = ratings, aes_string(x = column, y = "y.pos"), shape = 21, size = 5.5, color = "#3E4C59", fill = point_fill),
      geom_text(data = ratings, aes_string(x = column, y = "y.pos", label = column), family = "Roboto Condensed", color = text_col, size = 8) 
  )
}

p1 <- ggplot() +
  geom_text(aes(x = 8.9, y = 8:2, label = c("R", "A", "T","I", "N", "G", "S")), color = "#C5D3D9", family = "Bebas Neue", size = 20) +
  geom_segment(data = ratings, aes(x = min_rating, xend = max_rating, y = y.pos, yend = y.pos), color = "#3E4C59", linewidth = 0.25) +
  add_point_text(column = "min_rating", point_fill = "#F0F0F2", text_col = "#3E4C59") +
  add_point_text(column = "max_rating", point_fill = "#F0F0F2", text_col = "#3E4C59") +
  add_point_text(column = "avg_rating", point_fill = "#3E4C59", text_col = "#F0F0F2") +
   geom_text(data = ratings, aes(x = 6.8, y = y.pos, label = paste0("S", season)), family = "Bebas Neue", color = "#C5D3D9", size = 15, hjust = 0.5) +
  geom_text(data = filter(ratings, season == 1), aes(x = min_rating, y = 9.8, label = "min"), family = "Roboto Condensed", color = "#3E4C59", size = 8) +
  geom_text(data = filter(ratings, season == 1), aes(x = avg_rating, y = 9.8, label = "average"), family = "Roboto Condensed", color = "#3E4C59", size = 8) +
  geom_text(data = filter(ratings, season == 1), aes(x = max_rating, y = 9.8, label = "max"), family = "Roboto Condensed", color = "#3E4C59", size = 8) +
  ylim(c(-0.25, 10.25)) +
  theme_void() +
  theme(panel.background = element_rect(fill = "#F0F0F2", color = NA))
```

Days lasted
```{r}
p2 <- ggplot() +
  geom_segment(data = segment_vals,
               aes(x = day, xend = day, y = y.pos - 0.25, yend = y.pos + 0.25, color = day), linewidth = 1, show.legend = FALSE) +
  scale_color_gradient2(low = "#C5D3D9", mid = "#A4B8BF", high = "#697F8C", midpoint = 50) +
   geom_point(data = all_days, aes(x = days_lasted, y = y.pos + 0.35), shape = 21, size = 1, color = "#3E4C59", stroke = 0.3, fill = "transparent") +
  geom_text(data = all_days, aes(x = -2, y = y.pos, label = paste0("S", season)), family = "Bebas Neue", color = "#C5D3D9", size = 15, hjust = 1) + 
  geom_text(aes(x = 95, y = 8:5, label = c("D", "A", "Y", "S")), color = "#C5D3D9", family = "Bebas Neue", size = 20) + 
    scale_x_continuous(breaks = c(25, 50, 75, 100)) +
  labs(caption = "Number of days each season lasted.\nDots indicate the days when candidates quit or won.") +
  theme_minimal() +
  theme(panel.background = element_rect(fill = "#F0F0F2", color = NA),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(color = "#3E4C59", linetype = "dotted", linewidth = 0.2),
        axis.title = element_blank(),
        axis.text.x = element_text(family = "Roboto Condensed", color = "#3E4C59", size = 25),
        axis.text.y = element_blank(),
        plot.caption = element_text(family = "Roboto Condensed", color = "#3E4C59", size = 30, hjust = 0.5, lineheight = 0.25))
```

Items - top 5 items by season
https://twitter.com/birusod/status/1618855900193914880
https://github.com/davidsjoberg/ggbump/wiki/My-year-on-Spotify

```{r, warning=FALSE}
set.seed(42)
custom_palette <- c(RColorBrewer::brewer.pal(9, "Set1"),
                    RColorBrewer::brewer.pal(5, "Dark2")) %>% 
  sample(n_distinct(items$item))

p3 <- items %>% 
  ggplot(aes(order, rank, color = item, group = item)) +
  geom_bump(smooth = 15, size = 2, alpha = 0.2) +
  scale_y_reverse() +
  geom_bump(data = items %>% filter(rank <= 5), 
            aes(order, rank, group = group, color = item), 
            smooth = 15, size = 2, inherit.aes = F, alpha = 0.8) +
  geom_point(data = items %>% filter(d_first_top5 == 1),
             aes(x = order - .2), size = 5, alpha = 0.8) +
  geom_segment(data = items %>% filter(rank <=5),
               aes(x = order - .2, xend = order + .2, y = rank, yend = rank), size = 2, lineend = "round", alpha = 0.8) +
  scale_x_continuous(breaks = items$order %>% unique() %>% sort(),
                     labels = items %>% distinct(order, season) %>% arrange(order) %>% pull(season), 
                     expand = expand_scale(mult = .1)) +
  geom_text(data = items %>% filter(d_first_top5 == 1),
            aes(label = item, x = order-.2),
            color = "#3E4C59",
            nudge_y = .43,
            nudge_x = -.05,
            size = 9,
            fontface = 2,
            hjust = 0, 
            family = "Roboto Condensed") +
  geom_text(data = items %>% filter(order == max(order)),
            aes(label = item),
            color = "#3E4C59",
            nudge_x = .31,
            hjust = 0,
            size = 9,
            fontface = 2, 
            family = "Roboto Condensed") +
  theme_void() +
  theme(legend.position = "none",
        plot.caption = element_text(family = "Roboto Condensed", color = "#3E4C59", size = 30, hjust = 0.5, lineheight = 0.25),
        panel.background = element_rect(fill = "#F0F0F2", color = NA)) +
  labs(x = NULL,
       caption = "Top 5 items spanning the seasons") +
  scale_colour_manual(values = custom_palette) +
  geom_text(data = tibble(x = 0, y = 1:5), aes(x = x, y = y, label = y), color = "#C5D3D9", family = "Bebas Neue", size = 20, inherit.aes = F) +
  geom_text(data = items %>% distinct(season), aes(x = 1:9, y = 6.5, label = paste0("S", season)), inherit.aes = F, family = "Bebas Neue", color = "#C5D3D9", size = 15, hjust = 1) +
  xlim(c(0, 10))
```

# Combine plots into one big plot
```{r}
library(patchwork)
p <- m + p1 + p2 + p3 +
  plot_layout(ncol = 2, widths = c(1, 1, 1, 1)) +
  plot_annotation(
    title = "Alone -  an American survival competition TV series",
    caption = "Graphic by Rahim Jutha | Source : {alone} package by Dan Oehm",
    theme = theme(
      plot.background = element_rect(fill = "#F0F0F2", color = NA),
      panel.background = element_rect(fill = "#F0F0F2", color = NA), 
      plot.title = element_text(family = "Bebas Neue", color = "#3E4C59", size = 125, hjust = 0.5, margin = margin(t = 10, b = 10)),
      plot.caption = element_text(family = "Roboto Condensed", color = "#3E4C59", size = 30, hjust = 0.5)))

ggsave("alone.png", p, dpi = 320, width = 12, height = 6)
```